\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses with Citizen Science

### 2015-05-10 doc/sw-system.tex
Addressed all Christine's comments, fixes #205

\title[\SW]
{\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses}

@@

Strong gravitational lensing -- the formation of multiple, magnified images of
background objects due to the deflection of light by  massive foreground
objects -- is a very powerful astrophysical tool, enabling a wide range of
science projects. The image separations and distortions provide information
about the mass distribution in the lens \citep[e.g.][]{AugerEtal2010,
SonnenfeldEtal2012,MoreEtal2012,SonnenfeldEtal2013}, including on sub-galactic scales
\citep[e.g.][]{Dalal+Kochanek2002,VegettiEtal2010,HezavehEtal2013}. Any strong
lens can provide magnification of a factor of 10 or more, providing a deeper,
higher resolution view of the distant universe through these ``cosmic
telescopes'' \citep[e.g.][]{StarkEtal2008,NewtonEtal2011}. Lensed quasars
enable cosmography via the time delays between the multiple images'
lightcurves \citep[e.g.][]{TewesEtal2013,SuyuEtal2013}, and study of the
accretion disk itself through the microlensing effect
\citep[e.g.][]{PoindexterEtal2008}. All of these investigations would
benefit from being able to draw from a larger and/or more diverse sample of lenses.

@@

Because they are rare, strong lenses are expensive to find. The highest purity
searches to date have made use of relatively clean signals such as the
presence of emission or absorption features at two distinct redshifts in the
same optical spectrum \citep[e.g.][]{BoltonEtal2004}, or the strong
``magnification bias'' towards detecting strongly-lensed sources in the sub-mm
waveband \citep[e.g.][]{NegrelloEtal2010}. Such searches have to yield pure samples,
because they require expensive high resolution imaging follow-up; consequently
they have so far produced yields of only tens to hundreds of lenses. An alternative
approach is to search images already of sufficiently high resolution and colour
contrast, and confirm the systems as gravitational lenses by modeling the
survey data themselves \citep[][]{MarshallEtal2009}. Several square degrees of
HST images have been searched, yielding several tens of galaxy-scale lenses
\citep[e.g.][]{MoustakasEtal2007,FaureEtal2008,Jackson2008,MoreEtal2011,
PawaseEtal2014}. Similarly, searches of over a hundred square degrees of CFHT
Legacy Survey (\cfhtls) ground-based imaging, also with sub-arcsecond image quality,
have revealed a smaller number of wider image separation group-scale systems
\citep[e.g.][]{CabanacEtal2007,MoreEtal2012}. Detecting galaxy-scale lenses
from the ground is difficult, but feasible albeit with lower efficiency and requiring
HST or spectroscopic follow-up to confirm the candidates as lenses
\citep[e.g.][]{GavazziEtal2014}.

@@

Because we include a training set of images in amongst the survey ``test''
images, \sw can be viewed as a {\it supervised learning} system. Consequently,
we should expect to find lens candidates that look like the sims that we put in
the training set, and so the design of this training set is quite important. The
training images must look realistic, in two ways. First, apart from the
simulated lenses they contain, they must look like all the other survey images --
otherwise they could be identified as training images without the volunteer
learning anything about what lenses look like. For this reason, we make the sims
by {\it adding in} simulated lensed features to real images drawn from the
survey, and we make the duds by simply inspecting and choosing from a small
sample of randomly-selected survey images. It also means that each dataset
presented for inspection in the \sw interface needs to come with its own
training set.

@@

At this point, we applied several cuts to the simulated lens pool, in order to
reject sims that were likely to be difficult to spot. While this ensured that
the educational value of the sims when feedback was given, it means that we
should not necessarily expect to detect real lens candidates that would not pass
these cuts (although human classifiers do possess the key advantage of  being
able to go beyond their training\footnote{Any real lens candidates we find with
properties outside the ranges of the training set will be of particular
interest, being a product of the {\it imagination} of the crowd in extrapolating
their understanding of gravitational lenses into a regime that they have not
been trained for.  \citet{GeachEtal2015}, for example, report the discovery of a
red lensed arc  by volunteers who had only been trained primarily on blue
examples. We investigate this aspect of the \SW approach in \PaperTwo.}).
In the present project, the most important selection cuts we made were to only
keep sims with a) second brightest image brighter that $i = 23$ and b) total
combined image magnitude fainter than 19--20 (\PaperTwo, Table~1). The
combination of these two cuts resulted in a sample of visible lensed image
configurations.

Volunteers were initially shown training images at
a mean frequency of two in five. Subjects were drawn at random from a pool
consisting of (at first) 20\% sims, 20\% duds, and 60\% test images,
and such that no volunteer ever
saw a given image more than once. As the number of classifications made by the
volunteer increased, the training frequency was decreased from 40\% to
$2/(5\times2^{(\textrm{int}(N_c/20)+1)/2})$ ($\approx 30\%$ for the second 20
subjects, $20\%$ for the third 20 subjects, and so on).

@@

The four CFHT Legacy Survey\footnote{\texttt{http://www.cfht.hawaii.edu/Science/CFHTLS/}}
\citep[\cfhtls,][]{Gwyn2012} ``Wide'' fields cover a total of approximately
160 square degrees of sky (after taking into account tile overlaps).
With high and homogeneous image quality (the mean seeing in the $g$-band is
0.78''), and reaching limiting magnitudes of around 25 across the $ugriz$ filter
set, this survey has yielded several dozen new gravitational lenses on both
galaxy and group scales
\citep{GavazziEtal2014,SonnenfeldEtal13,CabanacEtal2007,MoreEtal2012}. The
quality of the data, combined with the presence of these comparison ``known
lens'' samples, makes this a natural choice against which to develop and test
the \SW system. The \cfhtls is also well-representative of the data quality
expected from several next-generation sky surveys, such as DES, KiDS, HSC and
LSST. We use the stacked images
from the final T0007 release taken from the Terapix
website\footnote{\texttt{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}} for this
work.

In order to investigate the completeness of the previous, semi-automated, lens
searches in the \cfhtls area, we designed a ``blind search.'' We divided the
\cfhtls pointings into some 430,000 equal size, overlapping tiles, approximately
82~arcsec on a side. We refer to these images as ``test images.'' The ``training
images'' were derived from a small subset of these, as  discussed above. In
future, larger area, projects we expect to implement a somewhat different
strategy of producing image tiles centred on particular pre-selected
``targets,'' which might make for a more efficient (if less complete) survey.
However, we do not expect the performance of citizen image inspectors to change
significantly between these strategies: to first order, both strategies require
the inspectors to learn what lenses look like, and then search the presented
images for similar features.

@@

While the \SW web app is live, and classifications are being made, we perform an
online analysis of the classifications,  updating a probabilistic model of every
(anonymous) volunteer's data, and also updating the posterior probability that
each subject (in both the training and test sets) contains a lens, on a daily
basis. This gives us a dynamic estimate of the posterior probability for  any
given  subject being a lens, given all classifications of it to date. Assigning
thresholds in this lens probability then allows us to make good decisions about
whether or not to retire a subject from the system, in order to focus attention
on new images.

@@

How should we model these compressed data? The circumstances of each
classification are quite complex, as are the human classifiers in general: the
volunteers learn more about the problem as they go, but also inevitably make
occasional mistakes (especially when classifying at high speed).
To cope with this uncertainty, we assign a
simple software {\it agent} to partner each volunteer. The agent's task is to
interpret their volunteer's classification data as best it can, using a model
that makes a number of necessary approximations. These interpretations will
then include uncertainty arising as a result of the volunteer's efforts and
also the agent's approximations, but they will have two important redeeming
features. First, the interpretations will be quantitative (where before they
were qualititative),  and thus will be useful in decision-making. Second, the
agent will be able to predict, using its model, the probability of a test
subject being a $\LENS$, given both its volunteer's classification, and its
volunteer's experience. In this section we describe how these agents work,
and other aspects of the \SW analysis pipeline (SWAP).

@@

Now, when this volunteer views a test subject,
it is this confusion matrix that will allow their agent to update the
probability of that test subject being a $\LENS$. Let us suppose that
this subject has never been seen before: the agent assigns a
prior probability that it is (or contains) a lens is
\be
  \pr(\LENS) = p_0
\ee
where we have to assign a value for $p_0$. In the \cfhtls, we might expect
something like 100 lenses in 430,000 images, so $p_0 = 2\times10^{-4}$
is a reasonable estimate. The volunteer then makes a classification $C_k$
($= \saidLENS$ or $\saidNOT$).
We can apply Bayes' Theorem to derive how the agent should
update this prior probability into a posterior one using this new information:
\begin{align}
  \label{eq:app:first}
  & \pr(\LENS|C_k,\trainingk) = \\
  & \frac{\pr(C_k|\LENS,\trainingk)\cdot\pr(\LENS)}
{\left[ \pr(C_k|\LENS,\trainingk)\cdot\pr(\LENS) + \pr(C_k|\NOT,\trainingk)\cdot\pr(\NOT) \right]},
  \notag
\end{align}
which can be evaluated numerically using the elements of the confusion
matrix.

@@

Suppose we have a volunteer who is always right about the true
nature of a training subject.
Their agent's confusion matrix would be
\be
  \CM^{\rm perfect} =
  \begin{bmatrix}
    0.0 & 1.0 \\
    1.0 & 0.0
  \end{bmatrix}.
\ee
On being given a fresh subject that actually is a $\LENS$, this hypothetical
volunteer would submit $C = \saidLENS$.  Their agent would then calculate the
posterior probability for the subject being a $LENS$ to be
\begin{align}
  \pr(\LENS|\saidLENS,\trainingk) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0,
\end{align}
as we might expect for such a {\it perfect} classifier.  Meanwhile, a
hypothetical volunteer who (for some reason) wilfully always submits the wrong
classification would have an agent with the column-swapped confusion matrix
\be
  \CM^{\rm obtuse} =
  \begin{bmatrix}
    1.0 & 0.0 \\
    0.0 & 1.0
  \end{bmatrix},
\ee
and would submit $C = \saidNOT$ for this subject. However, such a volunteer
would nevertheless be submitting useful information, since given the above
confusion matrix, their agent would calculate
\begin{align}
  \pr(\LENS|\saidNOT,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0.
\end{align}
{\it Obtuse} classifiers turn out to be as helpful as {\it perfect} ones, because
the agents know to trust the opposite of their classifications.

@@

Suppose the $k+1^{\rm th}$ volunteer now submits a classification, on the same
subject just classified by the $k^{\rm th}$ volunteer. We can generalise
\Eref{eq:app:first} by replacing the prior probability with the current
posterior probability:
\begin{align}
  \label{eq:app:update}
  \pr(\LENS & |C_{k+1},\training_{k+1},\data) = \\
  & \frac{1}{Z} \pr(C_{k+1}|\LENS,\training_{k+1}) \cdot \pr(\LENS|\data) \\ \notag
{\rm where}\;\; Z = & \pr(C_{k+1}|\LENS,\training_{k+1})\cdot\pr(\LENS|\data) \\ \notag
      & + \pr(C_{k+1}|\NOT,\training_{k+1})\cdot\pr(\NOT|\data), \notag
\end{align}
and $\data = \{C_k,\trainingk\}$ is the set of all previous
classifications, and the set of training subjects seen by each of those
volunteers.
$\pr(\LENS|\data)$ is the fundamental property of each test subject that
we are trying to infer. We track $\pr(\LENS|\data)$ as a function of time,
and by comparing it to a lower or upper thresholds, make decisions about
whether to retire the subject from the classification interface or
promote it in \Talk, respectively.

@@

The probabilistic model described above does not need to be implemented as an
online inference. Indeed, it might be more appropriate to perform the inference
of all agent confusion matrix elements and subject probabilities simultaneously.
Such a joint analysis would implement in full the agents' basic model assumption
that their volunteers have innate and unchanging talent for lens spotting, that
is parameterised by two constant confusion matrix elements which simply need to
be inferred given the data.

@@

\Fref{fig:swap:agent-probabilities} shows the confusion matrix elements of 200
randomly-selected agents, as they were on a particular day during the
\StageOne online analysis. Many volunteers classify  only a small number of images, and
so their agents' confusion matrix elements have remained close to their initial values
of (0.5,0.5). As more images are classified (shown by the yellow point size),
the agents' matrix elements tend to move towards higher values, partly as the
volunteers attain greater skill levels (green point sizes, see
\Sref{sec:results} below) but mostly as
the agent learns more about them. In this upper
right-hand quadrant, the agents perceive their volunteers to be ``astute.''

@@

For comparison we show the results of an analysis where the classifications of
training images were ignored, and the agents' confusion matrix elements
were instead all simply assigned initial values of 0.75
which then remained constant. This set-up emulates a
very simple unweighted voting scheme, where all classifications are treated
equally.  In this case, the TPR remains under 80\% in \StageOne and 60\% in
\StageTwo, thus illustrating the benefit of including training images and
allowing the agents to learn via their Bayesian updates.
When the agents are allowed to update their confusion matrices, the choice of
initial confusion matrix is not very important: the same 0.75 initial values
applied to normal, learning agents resulted in only a slightly lower TPR than
the default case.

The dot-dashed curves show the results from the offline analysis. At \StageOne the
results are very similar to the online version that was actually run (solid
line). However, at \StageTwo there is marginal evidence of there being greater
benefit to doing the analysis offline (\Sref{sec:swap:offline}).
Over 85\% TPR is achieved at zero FPR in the offline analysis , while if one is
willing to accept a false positive rate of 5\%, the true positive rate rises to
over 95\%, showing that some of the sims that were missed in the online analysis
may be being recovered by doing the analysis offline. (The same is true at
\StageOne, but to a lesser extent.) One interpretation for this result (if it holds up)
would be that the offline analysis, by using each agent's entire history, is
making less noisy probability estimates. This could be consistent with
other citizen science/crowdsourcing projects where,
in the absence of high quality information about classifiers,
sophisticated strategies tend to under-perform naive but simpler
ones \citep{Waterhouse} (although we note that this rule of thumb seems not to extend
to simple voting in this case!).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{minipage}{0.48\linewidth}
  \centering\includegraphics[width=\linewidth]{stage1_ROC.pdf}
\end{minipage}\hfill
\begin{minipage}{0.48\linewidth}
  \centering\includegraphics[width=\linewidth]{stage2_ROC.pdf}
\end{minipage}
\caption{Receiver operating characteristic curves for the \SW system, using
the \cfhtls training set. Left: \StageOne, right: \StageTwo. Insets show a
magnified view of the top lefthand corner of each plot. Linestyles illustrate
the agent properties (initial confusion matrix elements, learning or not)
discussed in the text, as well as the difference between the online and offline
analyses. The \StageTwo sample was defined using the online \StageOne results
(solid blue curve), while the offline analysis was chosen for the final
\StageTwo results (dot-dashed orange curve).
Lens probability varies along the curves: the stars indicate the point where
the value of this quantity is 0.95.}
\label{fig:results:sample:roc}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

To investigate the properties of the \sw crowd, as characterised by their
agents, we define the following quantities and plot their one and 2-D marginal
distributions in
\Fref{fig:crowd:cornerplot}. This figure only shows the \StageOne agents for clarity,
but the trends we found the same trends in \StageTwo.
%
\begin{description}
%
\item{\noindent\bf ``Effort:''} The number of test images, $\effort$, classified
by a volunteer. In \StageOne, the mean effort per agent was 263; in the shorter
\StageTwo it was 81.
%
\item{\noindent\bf ``Experience:''} The number of training images, $\experience$,
classified by a volunteer. In \StageOne, the mean experience per agent was 29;
in \StageTwo (where the training image frequency was set higher) it was 34.
%
\item{\noindent\bf ``Skill:''} The expectation value of the information gained
per classification (in bits) by that volunteer,  $\skill$, for subjects which
have lens probability 0.5 (\Aref{appendix:swap}). Random classifiers have
$\skill = 0.0$~bits, while perfect classifiers have $\skill = 1.0$~bit.
All agents start
with $\skill = 0.0$~bits. The skill of  an agent increases as training subjects are
classified, and the agent's estimates of its confusion matrix elements improve.
In \StageOne, the mean skill per agent was 0.04 bits; in  \StageTwo it was
0.05~bits.
%
\item{\noindent\bf ``Contribution:''} This is the integrated skill over a volunteer's
{\it test subject} classification history, a quantity representative of the
total contribution to the project made by that volunteer (see \Aref{appendix:swap} for
more discussion of this quantity). The classifications of training images
allow us to estimate the skill, while the classification of test images determines
contribution.  In \StageOne, the mean  contribution per agent was
34.9 bits; in \StageTwo it was 33.5.
%
\item{\noindent\bf ``Information:''} The total information $\information$
generated by the agent during the volunteer's classification activity. This
quantity depends on the value of each subject's lens probability when that
subject was presented to the volunteer (simply because information gain is
defined in terms of posterior relative to prior probabilities,
\Aref{appendix:swap}).  Because of this, we expect this raw quantity to be noisier
than the ``contribution'' defined above. The lower righthand panel of
\Fref{fig:crowd:cornerplot} confirms this: there is a strong, linear correlation
between contribution and information gain, but at a given contribution, the
distribution of information gain has a tail at high values, corresponding to
volunteers that were lucky in the number of high probability subjects that they
happened to be shown (the information gain per classification is maximized when the
subject probability is 0.5, while most subjects have probabilities closer to or less than the prior).
In most of the following discussion we use the contribution when characterizing
crowd participation, but we include
the information gain for completeness.
%
\end{description}

>>>>

\title[\SW]
{\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses with Citizen Science}

@@

Strong gravitational lensing -- the formation of multiple, magnified images of
background objects due to the deflection of light by  massive foreground
objects -- is a very powerful astrophysical tool, enabling a wide range of
science projects. The image separations and distortions provide information
about the mass distribution in the lens \citep[e.g.][]{AugerEtal2010,
SonnenfeldEtal2012,MoreEtal2012,SonnenfeldEtal2015}, including on sub-galactic scales
\citep[e.g.][]{Dalal+Kochanek2002,VegettiEtal2010,HezavehEtal2013}. Any strong
lens can provide magnification of a factor of 10 or more, providing a deeper,
higher resolution view of the distant universe through these ``cosmic
telescopes'' \citep[e.g.][]{StarkEtal2008,NewtonEtal2011}. Lensed quasars
enable cosmography via the time delays between the multiple images'
lightcurves \citep[e.g.][]{TewesEtal2013,SuyuEtal2013}, and study of the
accretion disk itself through the microlensing effect
\citep[e.g.][]{PoindexterEtal2008}. All of these investigations would
benefit from being able to draw from a larger and/or more diverse sample of lenses.

@@

Strong lenses are expensive to find, because they are rare. The highest purity
searches to date have made use of relatively clean signals such as the
presence of emission or absorption features at two distinct redshifts in the
same optical spectrum \citep[e.g.][]{BoltonEtal2004}, or the strong
``magnification bias'' towards detecting strongly-lensed sources in the sub-mm
waveband \citep[e.g.][]{NegrelloEtal2010}. Such searches have to yield pure samples,
because they require expensive high resolution imaging follow-up; consequently
they have so far produced yields of only tens to hundreds of lenses. An alternative
approach is to search images already of sufficiently high resolution and colour
contrast, and confirm the systems as gravitational lenses by modeling the
survey data themselves \citep[][]{MarshallEtal2009}. Several square degrees of
HST images have been searched, yielding several tens of galaxy-scale lenses
\citep[e.g.][]{MoustakasEtal2007,FaureEtal2008,Jackson2008,MoreEtal2011,
PawaseEtal2014}. Similarly, searches of over a hundred square degrees of CFHT
Legacy Survey (\cfhtls) ground-based imaging, also with sub-arcsecond image quality,
have revealed a smaller number of wider image separation group-scale systems
\citep[e.g.][]{CabanacEtal2007,MoreEtal2012}. Detecting galaxy-scale lenses
from the ground is difficult, but feasible albeit with lower efficiency and requiring
HST or spectroscopic follow-up to confirm the candidates as lenses
\citep[e.g.][]{GavazziEtal2014}.

@@

\sw can be viewed as a {\it supervised learning} system,
because we include a training set of images in amongst the survey ``test''
images. Consequently,
we should expect to find lens candidates that look like the sims that we put in
the training set, and so the design of this training set is quite important. The
training images must look realistic, in two ways. First, apart from the
simulated lenses they contain, they must look like all the other survey images --
otherwise they could be identified as training images without the volunteer
learning anything about what lenses look like. For this reason, we make the sims
by {\it adding in} simulated lensed features to real images drawn from the
survey, and we make the duds by simply inspecting and choosing from a small
sample of randomly-selected survey images. It also means that each dataset
presented for inspection in the \sw interface needs to come with its own
training set.

@@

At this point, we applied several cuts to the simulated lens pool, in order to
reject sims that were likely to be difficult to spot. While this ensured
that the sims had high educational value, it means that we
should not necessarily expect to be able to
detect real lens candidates that would not pass these cuts.
However, human classifiers do possess the key advantage of being
able to {\it imagine} lens systems beyond what they were shown during their training.
Indeed, any real lens candidates we find with
properties outside the ranges of the training set could be of particular
interest, if the training set is indeed completely representative of lenses that
we already know about.
\citet{GeachEtal2015}, for example, report the discovery of a
red lensed arc  by volunteers who had only been trained primarily on blue
examples. We investigate this aspect of the \SW approach further in \PaperTwo.
In the present project, the most important selection cuts we made were to only
keep sims with a) second brightest image having $i < 23$ and b) total
combined image magnitude fainter than 19--20 (\PaperTwo, Table~1). The
combination of these two reductions resulted in a sample of visible lensed image
configurations.

Volunteers were initially shown training images at
a mean frequency of two in five. Subjects were drawn at random from a pool
consisting of (at first) 20\% sims, 20\% duds, and 60\% test images,
and such that no volunteer ever
saw a given image more than once. As the number of classifications made by the
volunteer increased, the training frequency was decreased from 40\% to
$2/(5\times2^{(\textrm{int}(N_c/20)+1)/2})$, $\approx 30\%$ for the second 20
subjects, $20\%$ for the third 20 subjects, and the minimum rate of 10\%
after that. We did not reduce the training frequency to zero because we wanted
to ensure that the inspectors remained alert.

@@

The four CFHT Legacy Survey\footnote{\texttt{http://www.cfht.hawaii.edu/Science/CFHTLS/}}
\citep[\cfhtls,][]{Gwyn2012} ``Wide'' fields cover a total of approximately
160 square degrees of sky (after taking into account tile overlaps).
With high and homogeneous image quality (the mean seeing in the $g$-band is
0.78''), and reaching limiting magnitudes of around 25 across the $ugriz$ filter
set, this survey has yielded several dozen new gravitational lenses on both
galaxy and group scales
\citep{GavazziEtal2014,SonnenfeldEtal2013a,CabanacEtal2007,MoreEtal2012}. The
quality of the data, combined with the presence of these comparison ``known
lens'' samples, makes this a natural choice against which to develop and test
the \SW system. The \cfhtls is also well-representative of the data quality
expected from several next-generation sky surveys, such as DES, KiDS, HSC and
LSST. We use the stacked images
from the final T0007 release taken from the Terapix
website\footnote{\texttt{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}} for this
work.

In order to investigate the completeness of the previous semi-automated lens
searches in the \cfhtls area, we designed a ``blind search'' as follows.
We divided the
\cfhtls pointings into some 430,000 equal size, overlapping tiles, approximately
82~arcsec on a side. We refer to these images as ``test images.'' The ``training
images'' were derived from a small subset of these, as  discussed above. In
future, larger area, projects we expect to implement a somewhat different
strategy of producing image tiles centred on particular pre-selected
``targets,'' which might make for a more efficient (if less complete) survey.
However, we do not expect the performance of citizen image inspectors to change
significantly between these strategies: to first order, both strategies require
the inspectors to learn what lenses look like, and then search the presented
images for similar features.

@@

While the \SW web app is live, and classifications are being made, we perform an
online analysis of the classifications,  updating a probabilistic model of every
anonymous volunteer's data, and also updating the posterior probability that
each subject (in both the training and test sets) contains a lens, on a daily
basis. This gives us a dynamic estimate of the posterior probability for  any
given  subject being a lens, given all classifications of it to date. Assigning
thresholds in this lens probability then allows us to make good decisions about
whether or not to retire a subject from the system, in order to focus attention
on new images.

@@

How should we model these compressed data? The circumstances of each
classification are quite complex, as are the human classifiers in general: the
volunteers learn more about the problem as they go, but also inevitably make
occasional mistakes (especially when classifying at high speed).
To cope with this uncertainty, we assign a
simple software {\it agent} to partner each volunteer. The agent's task is to
interpret their volunteer's classification data as best it can, using a model
that makes a number of necessary approximations. These interpretations will
then include uncertainty arising as a result of the volunteer's efforts and
also the agent's approximations.
The agent will be able to predict, using its model, the probability of a test
subject being a $\LENS$ given both its volunteer's classification and its
volunteer's past experience. In this section we describe how these agents work,
and other aspects of the \SW analysis pipeline (SWAP).

@@

Now, when this volunteer views a test subject,
it is this confusion matrix that will allow their software agent to update the
probability of that test subject being a $\LENS$. Let us suppose that
this subject has never been seen before: the agent assigns a
prior probability that it is (or contains) a lens is
\be
  \pr(\LENS) = p_0
\ee
where we have to assign a value for $p_0$. In the \cfhtls, we might expect
something like 100 lenses in 430,000 images, so $p_0 = 2\times10^{-4}$
is a reasonable estimate. The volunteer then makes a classification $C_k$
($= \saidLENS$ or $\saidNOT$).
We can apply Bayes' Theorem to derive how the agent should
update this prior probability into a posterior one using this new information:
\begin{align}
  \label{eq:app:first}
  & \pr(\LENS|C_k,\trainingk) = \\
  & \frac{\pr(C_k|\LENS,\trainingk)\cdot\pr(\LENS)}
{\left[ \pr(C_k|\LENS,\trainingk)\cdot\pr(\LENS) + \pr(C_k|\NOT,\trainingk)\cdot\pr(\NOT) \right]},
  \notag
\end{align}
which can be evaluated numerically using the elements of the confusion
matrix.

@@

Suppose we have a volunteer who is always right about the true
nature of a training subject.
Their agent's confusion matrix would be
\be
  \CM^{\rm perfect} =
  \begin{bmatrix}
    0.0 & 1.0 \\
    1.0 & 0.0
  \end{bmatrix}.
\ee
On being given a fresh subject that actually is a $\LENS$, this hypothetical
volunteer would submit $C = \saidLENS$.  Their agent would then calculate the
posterior probability for the subject being a $LENS$ to be
\begin{align}
  \pr(\LENS|\saidLENS,\trainingk) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0,
\end{align}
as we might expect for such a {\it perfect} classifier.  Meanwhile, a
hypothetical volunteer who (for some reason) wilfully always submits the wrong
classification would have an agent with the column-swapped confusion matrix
\be
  \CM^{\rm obtuse} =
  \begin{bmatrix}
    1.0 & 0.0 \\
    0.0 & 1.0
  \end{bmatrix},
\ee
and would submit $C = \saidNOT$ for this subject. However, such a volunteer
would nevertheless be submitting useful information, since given the above
confusion matrix, their software agent would calculate
\begin{align}
  \pr(\LENS|\saidNOT,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0.
\end{align}
{\it Obtuse} classifiers turn out to be as helpful as {\it perfect} ones, because
the agents know to trust the opposite of their classifications.

@@

Suppose the $k+1^{\rm th}$ volunteer now submits a classification, on the same
subject just classified by the $k^{\rm th}$ volunteer. We can generalise
\Eref{eq:app:first} by replacing the prior probability with the current
posterior probability:
\begin{align}
  \label{eq:app:update}
  \pr(\LENS & |C_{k+1},\training_{k+1},\data) = \\
  & \frac{1}{Z} \pr(C_{k+1}|\LENS,\training_{k+1}) \cdot \pr(\LENS|\data) \\ \notag
{\rm where}\;\; Z = & \pr(C_{k+1}|\LENS,\training_{k+1})\cdot\pr(\LENS|\data) \\ \notag
      & + \pr(C_{k+1}|\NOT,\training_{k+1})\cdot\pr(\NOT|\data), \notag
\end{align}
and $\data = \{C_k,\trainingk\}$ is the set of all previous
classifications, and the set of training subjects seen by each of those
volunteers.
$\pr(\LENS|\data)$ is the fundamental property of each test subject that
we are trying to infer. We track $\pr(\LENS|\data)$ as a function of time,
and by comparing it to a lower or upper threshold, make decisions about
whether to retire the subject from the classification interface, or
flag it for further study, respectively.

@@

The probabilistic model described above does not need to be implemented as an
online inference. Indeed, it might be more appropriate to perform the inference
of all agent confusion matrix elements and subject probabilities simultaneously.
Such a joint analysis would implement in full the software agents' basic model assumption
that their volunteers have innate and unchanging talent for lens spotting, that
is parameterised by two constant confusion matrix elements which simply need to
be inferred given the data.

@@

\Fref{fig:swap:agent-probabilities} shows the confusion matrix elements of 200
randomly-selected agents, as they were on a particular day during the
\StageOne online analysis. Many volunteers classify  only a small number of images, and
so their software agents' confusion matrix elements have remained close to their initial values
of (0.5,0.5). As more images are classified (shown by the yellow point size),
the agents' matrix elements tend to move towards higher values, partly as the
volunteers attain greater skill levels (green point sizes, see
\Sref{sec:results} below) but mostly as
the agent updates its confusion matrix. In this upper
right-hand quadrant, the agents perceive their volunteers to be ``astute.''

@@

For comparison we show the results of an analysis where the classifications of
training images were ignored, and the agents' confusion matrix elements
were instead all simply assigned initial values of 0.75
which then remained constant. This set-up emulates a
very simple unweighted voting scheme, where all classifications are treated
equally.  In this case, the TPR remains under 80\% in \StageOne and 60\% in
\StageTwo, thus illustrating the benefit of including training images and
allowing the agents to learn via their Bayesian updates.
When the software agents are allowed to update their confusion matrices, the choice of
initial confusion matrix is not very important: the same 0.75 initial values
applied to normal, learning agents resulted in only a slightly lower TPR than
the default case.

The dot-dashed curves show the results from the offline analysis. At \StageOne the
results are very similar to the online version that was actually run (solid
line). However, at \StageTwo there is marginal evidence of there being greater
benefit to doing the analysis offline (\Sref{sec:swap:offline}).
Over 85\% TPR is achieved at zero FPR in the offline analysis , while if one is
willing to accept a false positive rate of 5\%, the true positive rate rises to
over 95\%, showing that some of the sims that were missed in the online analysis
may be being recovered by doing the analysis offline. (The same is true at
\StageOne, but to a lesser extent.) One interpretation for this result (if it holds up)
would be that the offline analysis, by using each agent's entire history, is
making less noisy probability estimates. This could be consistent with
other citizen science/crowdsourcing projects where,
in the absence of high quality information about classifiers,
sophisticated strategies tend to under-perform naive but simpler
ones \citep[][although we note that this rule of thumb seems not to extend
to simple voting in this case!]{Waterhouse}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{minipage}{0.48\linewidth}
  \centering\includegraphics[width=\linewidth]{stage1_ROC.pdf}
\end{minipage}\hfill
\begin{minipage}{0.48\linewidth}
  \centering\includegraphics[width=\linewidth]{stage2_ROC.pdf}
\end{minipage}
\caption{Receiver operating characteristic curves for the \SW system, using
the \cfhtls training set. Left: \StageOne, right: \StageTwo. Insets show a
magnified view of the top lefthand corner of each plot. Linestyles illustrate
the software agent properties (initial confusion matrix elements, learning or not)
discussed in the text, as well as the difference between the online and offline
analyses. The \StageTwo sample was defined using the online \StageOne results
(solid blue curve), while the offline analysis was chosen for the final
\StageTwo results (dot-dashed orange curve).
Lens probability varies along the curves: the stars indicate the point where
the value of this quantity is 0.95.}
\label{fig:results:sample:roc}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

To investigate the properties of the \sw crowd, as characterised by their
agents, we define the following quantities and plot their one and 2-D marginal
distributions in
\Fref{fig:crowd:cornerplot}. This figure only shows the \StageOne agents for clarity,
but the trends we found the same trends in \StageTwo.
%
\begin{description}
%
\item{\noindent\bf ``Effort:''} The number of test images, $\effort$, classified
by a volunteer. In \StageOne, the mean effort per agent was 263; in the shorter
\StageTwo it was 81.
%
\item{\noindent\bf ``Experience:''} The number of training images, $\experience$,
classified by a volunteer. In \StageOne, the mean experience per agent was 29;
in \StageTwo (where the training image frequency was set higher) it was 34.
%
\item{\noindent\bf ``Skill:''} The expectation value of the information gained
per classification (in bits) by that volunteer,  $\skill$, for subjects which
have lens probability 0.5 (\Aref{appendix:swap}). Random classifiers have
$\skill = 0.0$~bits, while perfect classifiers have $\skill = 1.0$~bit.
All software agents start
with $\skill = 0.0$~bits. The skill of  an agent increases as training subjects are
classified, and the agent's estimates of its confusion matrix elements improve.
In \StageOne, the mean skill per agent was 0.04 bits; in  \StageTwo it was
0.05~bits.
%
\item{\noindent\bf ``Contribution:''} This is the integrated skill over a volunteer's
{\it test subject} classification history, a quantity representative of the
total contribution to the project made by that volunteer (see \Aref{appendix:swap} for
more discussion of this quantity). The classifications of training images
allow us to estimate the skill, while the classification of test images determines
contribution.  In \StageOne, the mean  contribution per agent was
34.9 bits; in \StageTwo it was 33.5.
%
\item{\noindent\bf ``Information:''} The total information $\information$
generated by the software agent during the volunteer's classification activity. This
quantity depends on the value of each subject's lens probability when that
subject was presented to the volunteer (simply because information gain is
defined in terms of posterior relative to prior probabilities,
\Aref{appendix:swap}).  As a result, we expect this raw quantity to be noisier
than the ``contribution'' defined above. The lower righthand panel of
\Fref{fig:crowd:cornerplot} confirms this: there is a strong, linear correlation
between contribution and information gain, but at a given contribution, the
distribution of information gain has a tail at high values, corresponding to
volunteers that were lucky in the number of high probability subjects that they
happened to be shown (the information gain per classification is maximized when the
subject probability is 0.5, while most subjects have probabilities closer to or less than the prior).
In most of the following discussion we use the contribution when characterizing
crowd participation, but we include
the information gain for completeness.
%
\end{description}
